Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-03-21 Delaneyetal2024.tex
more small changes

\begin{keypoints}
\item Short-term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, as is the case in subaerial systems.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Results suggest that significant subglacial sediment transport capacity can persist across a range of water discharge conditions, and the timing of peak events occurs under different hydrological conditions.
\end{keypoints}

@@

\section{Results}

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
    Gray (black) lines in a and e represent the subglacial channel size (water discharge).
    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
    Data are shown at $15$\unit{min} intervals.
    Arrows denote where variables peak in the subglacial channel before the subaerial one.
    Insets in f-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
    % Note that the x and y axes are different for \alpine{} and \icesheet{}.
  }
  \label{fig:model_outs}
\end{figure}

@@

In applying these models to different climate scenarios, water discharge measurements at the sub-hour timescale remain unavailable in many catchments.
The model outputs' variability in subglacial channels approaches that of subaerial ones  over $1-5$ days (Figure~\ref{fig:multi_run}), possibly resulting in parity between subaerial and subglacial channels over these periods.
This suggests that aggregating water discharge over daily to weekly periods could yield similar variations between subglacial and subaerial systems, although generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity there (Figure\,\ref{fig:model_outs}). 
Yet, water discharge over periods longer than days may omit key processes, such as the time of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).

@@

\item  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity variations;
  \vspace{.1cm}

\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of elevated water discharge in subglacial channels compared to subaerial ones.
  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects variations in timing.
\end{enumerate}

@@

\appendix
\TODO{match Friction factor in appendix}
\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier}
  \small % TODO remove again
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\
    && (\alpine{} or \icesheet{})\\
    \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (subglacial, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (subglacial, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (subglacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\

    Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\

@@

    \textbf{Parameters and Constants}  & & &\\
    Gravitational constant&$g$& $9.81$&$\mathrm{m\,s^{-2}}$\\
    Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
    Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
    Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5,\,16$ or $3$ & $\mathrm{(-)}$ \\
    Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
    Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
    Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\

>>>>

\begin{keypoints}
\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size, as in subaerial systems.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Significant subglacial sediment transport capacity can persist across a range of water discharges, impacting the  timing of peak events.
\end{keypoints}

@@

\section{Results}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Model outputs from simulations using the hydrographs in panels a and e for the scenarios \alpine{} (a-d) and \icesheet{} (e-h).
    Gray (black) lines in a and e represent the subglacial channel size (water discharge).
    Blue (orange) lines represent outputs from the subglacial (subaerial) channel.
    Data are shown at $15$\unit{min} intervals.
    Arrows denote where variables peak in the subglacial channel before the subaerial one.
    Insets in f-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
    % Note that the x and y axes are different for \alpine{} and \icesheet{}.
  }
  \label{fig:model_outs}
\end{figure}

@@

Water discharge measurements at the sub-hour timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of water aggregation time (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
However, generally steeper hydraulic gradients of glaciers may cause elevated sediment transport capacity in subglacial channels (Figure\,\ref{fig:model_outs}). 
Even so, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).

@@

\item
  water discharge variability in the two hydrographs impacts subglacial sediment transport capacity fluctuations;
  \vspace{.1cm}

\item the uneven response of sediment transport capacity to water discharge variations shows a different response and timing to periods of elevated water discharge in subglacial channels compared to subaerial ones.
  However, representing subglacial sediment transport capacity with either open-channel flow or pipe flow neglects differences in timing.
\end{enumerate}

@@

\appendix

\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a scenario from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a scenario from Leverett Glacier}
  \small % TODO remove again
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\
    && (\alpine{} or \icesheet{})\\
    \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (subglacial, subaerial)  & $v$, ($v_g,\,v_{f}$)& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (subglacial, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Channel depth (subaerial) & $H$&& $\mathrm{m}$\\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (subglacial, subaerial) & $w$, ($w_g,w_f$)&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Hydraulic gradient &$\Psi=\frac{\Delta h}{l}$&& $\mathrm{m\, m^{-1}}$\\
    Shear stress (subglacial, subaerial) & $\tau$, ($\tau_g,\,\tau_f$) && $\mathrm{Pa \, m^{-2}}$ \\
    Stream power & $\Omega$ && $\mathrm{ kg \, m\, s^{-3}}$ \\

@@

    \textbf{Parameters and Constants}  & & &\\
    Gravitational constant&$g$& $9.81$&$\mathrm{m\,s^{-2}}$\\
    Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
    Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
    Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
    Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
    Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
    Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
